Investigating factors that may influence COVID-19 survival

### 2022-01-19 manuscript.Rmd
Update manuscript.Rmd

---
title: Investigating factors that may influence COVID-19 survival
author: 36112985
bibliography: manuscript/bibliography.bib
output:
  rmarkdown::pdf_document:
    keep_tex: FALSE
    citation_package: "default"
    number_sections: FALSE
    extra_dependencies:
      - booktabs
urlcolor: orange
linkcolor: green
citecolor: red
header-includes:
  - \usepackage{color}
  - \usepackage{caption}
  - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \usepackage{fancyhdr}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{remark}{Remark}
  - \newtheorem{definition}{Definition}
  - \definecolor{darkgrey}{HTML}{4a4a4a}
  - \titleformat*{\section}{\Large\color{darkgrey}}
  - \pagestyle{fancy}
---

@@

library(magrittr)
library(latex2exp)
library(survival)
library(patchwork)
```

<!--- external functios --->
```{r}
sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())
sys.source(file = 'R/functions/StudyData.R', envir = knitr::knit_global())

sys.source(file = 'R/background/Numbers.R', envir = knitr::knit_global())
```

```{r}
data <- StudyData()
```

@@

\begin{table}[h!]
    \centering
    \renewcommand{\arraystretch}{1.25}
    \begin{tabular}{l r r r}
    & & & \\
    \textbf{In Hospital Mortality} & & & \\
    \hline
     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
    deceased & 31724 & 14049 & 347 \\
     & 0.688 & 0.305 & 0.008 \\
     & & & \\
    \textbf{Comorbidities} & & & \\
    \hline
     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
    asthma & 37376 & 5872 & 2872 \\
    & 0.81 & 0.13 & 0.06 \\
    mild liver disease & 42021 & 666 & 3433 \\
     & 0.91 & 0.01 & 0.07 \\
    renal disease & 35158 & 8056 & 2906 \\
     & 0.76 & 0.17 & 0.06 \\
    pulmonary & 35265 & 8078 & 2777 \\
     & 0.76 & 0.18 & 0.06 \\
    neurological disorder & 37227 & 5806 & 3087 \\
     & 0.81 & 0.13 & 0.07 \\
    liver mod severe &  41901 &  902 & 3317 \\
     & 0.91 & 0.02 & 0.07 \\
    malignant neoplasm & 38164 & 4734 & 3222 \\
     & 0.83 & 0.10 & 0.07 \\
     & & & \\
    \textbf{Demographics} & & & \\
    \hline
    \textbf{Age Group} &  \textbf{Male} & \textbf{Female} & \textbf{NA} \\
    80-89 & 7067 & 6197 & 31 \\
    & 0.153 & 0.134 & 0.001 \\
    70-79 & 6388 & 4445 & 19 \\
    & 0.139 & 0.096 & 0.000 \\
    60-69 & 4321 & 2601 & 9 \\
    & 0.094 & 0.056 & 0.000 \\
    50-59 & 3463 & 2151 & 9 \\
    & 0.075 & 0.047 & 0.000 \\
    90+ & 2085 & 2816 & 13 \\
    & 0.045 & 0.061 & 0.000 \\
    40-49 & 1846 & 1110 & 5 \\
    & 0.040 & 0.024 & 0.000 \\
    30-39 & 784 & 756 & 4 \\
    & 0.017 & 0.016 & 0.000 \\
    \hline
    \end{tabular}
\end{table}

@@

## Null Survival Curve

* Complete Case
* Imputation Case

## Uni-variate & Multivariate Analysis

* The table ... Cox

@@

@@

```{r}

```

@@

<!--- # References {#references .unnumbered} --->
# References

>>>>

---
title: Investigating factors that may influence COVID-19 survival
author: 36112985
bibliography: manuscript/bibliography.bib
output:
  rmarkdown::pdf_document:
    keep_tex: FALSE
    citation_package: "default"
    number_sections: FALSE
    extra_dependencies:
      - booktabs
urlcolor: orange
linkcolor: green
citecolor: red
header-includes:
  - \usepackage{color}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \usepackage{fancyhdr}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{remark}{Remark}
  - \newtheorem{definition}{Definition}
  - \definecolor{darkgrey}{HTML}{4a4a4a}
  - \titleformat*{\section}{\Large\color{darkgrey}}
  - \pagestyle{fancy}
---

@@

library(magrittr)
library(latex2exp)
library(survival)
library(patchwork)
library(kableExtra)
```

<!--- external functios --->
```{r}
sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())
sys.source(file = 'R/functions/ExtensiveStudyData.R', envir = knitr::knit_global())
sys.source(file = 'R/modelling/UnivariateAnalysis.R', envir = knitr::knit_global())
sys.source(file = 'R/modelling/MultivariateAnalysis.R', envir = knitr::knit_global())
sys.source(file = 'R/background/Numbers.R', envir = knitr::knit_global())
```

```{r}
dataframes <- ExtensiveStudyData(upload = TRUE)
data <- dataframes$data
training_ <- dataframes$training_
data_ <- dataframes$data_
```

@@

\begin{table}[h!]
    \scriptsize
    \centering
    \renewcommand{\arraystretch}{1.25}
    \begin{tabular}{l r r r}
    & & & \\
    \textbf{In Hospital Mortality} & & & \\
    \hline
     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
    deceased & 31724 & 14049 & 347 \\
     & 0.688 & 0.305 & 0.008 \\
     & & & \\
    \textbf{Comorbidities} & & & \\
    \hline
     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
    asthma & 37376 & 5872 & 2872 \\
    & 0.81 & 0.13 & 0.06 \\
    mild liver disease & 42021 & 666 & 3433 \\
     & 0.91 & 0.01 & 0.07 \\
    renal disease & 35158 & 8056 & 2906 \\
     & 0.76 & 0.17 & 0.06 \\
    pulmonary & 35265 & 8078 & 2777 \\
     & 0.76 & 0.18 & 0.06 \\
    neurological disorder & 37227 & 5806 & 3087 \\
     & 0.81 & 0.13 & 0.07 \\
    liver mod severe &  41901 &  902 & 3317 \\
     & 0.91 & 0.02 & 0.07 \\
    malignant neoplasm & 38164 & 4734 & 3222 \\
     & 0.83 & 0.10 & 0.07 \\
     & & & \\
    \textbf{Demographics} & & & \\
    \hline
    \textbf{Age Group} &  \textbf{Male} & \textbf{Female} & \textbf{NA} \\
    80-89 & 7067 & 6197 & 31 \\
    & 0.153 & 0.134 & 0.001 \\
    70-79 & 6388 & 4445 & 19 \\
    & 0.139 & 0.096 & 0.000 \\
    60-69 & 4321 & 2601 & 9 \\
    & 0.094 & 0.056 & 0.000 \\
    50-59 & 3463 & 2151 & 9 \\
    & 0.075 & 0.047 & 0.000 \\
    90+ & 2085 & 2816 & 13 \\
    & 0.045 & 0.061 & 0.000 \\
    40-49 & 1846 & 1110 & 5 \\
    & 0.040 & 0.024 & 0.000 \\
    30-39 & 784 & 756 & 4 \\
    & 0.017 & 0.016 & 0.000 \\
    \hline
    \end{tabular}
\end{table}

Correlation graph

@@

## Uni-variate & Multivariate Analysis: Training

The results of Table \@ref(tab:univariateTraining) summarise the separate effects of each potential prognostic factor on survival probability; investigated via Cox Proportional Hazard uni-variate analysis $\rightarrow$ coefficient: regression coefficients, hazard ratio & confidence interval: effect size, $p_{value}$: statistical significance in relation to survival.

In terms of age groups, the reference age group is *80 – 89*, and their $p_{value}$ scores suggests that **(a)** each age group, except *90+*, has a lower risk of death compared to age group *80 – 89*, and **(b)** the risk of death increases with age.  The sex $p_{value}$ suggests that females have a lower risk of death compared to males.  The $p_{value}$ scores for patients with an asthma, renal disease, pulmonary disease, or malignant neoplasm, diagnosis suggests that these patients have a higher risk of death compared to their respective disease/disorder free counterparts.

Alas, the proportionality tests invalidate most of the preceding observations; age group and malignant neoplasm have a score test value below $0.05$, the renal disease score test value is approximately $0.05$, and neurological disorder score is barely greater than $0.05$.  The multivariate analysis confirms the trend.

\footnotesize
```{r univariateTraining}
diagnostics <- UnivariateAnalysisTraining()

diagnostics %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    digits = 3,
    escape = FALSE,
    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
                                        'upper', '$p_{value}$', '${U}$' )),
    caption = 'Univariate Analysis', position = '!ht') %>%
    kableExtra::kable_styling(position = 'center') %>%
    kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
    kableExtra::add_footnote(c('CI: Confidence Interval',
                               'U: Proportionality score test'), notation = "number") %>%
    kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
    kableExtra::column_spec(column = 2, latex_column_spec = "r")
```
\normalsize

\footnotesize
```{r multivariateTraining}
diagnostics <- MultivariateAnalysisTraining()

diagnostics %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    digits = 3,
    escape = FALSE,
    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
                                        'upper', '$p_{value}$', '${U}$' )),
    caption = 'Multivariate Analysis', position = '!ht') %>%
  kableExtra::kable_styling(position = 'center') %>%
  kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
  kableExtra::add_footnote(c('CI: Confidence Interval',
                             'U: Proportionality score test'), notation = "number")  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "r")
```
\normalsize

## Uni-variate & Multivariate Analysis: All

Due to the failure of the Cox Proportional Hazard Modelling & Analysis strategy, ...

@@

\pagebreak

@@

```{r}
blob <- DiseaseNumbers(data = data)
knitr::kable(blob, caption = 'The disease frequencies', format = 'latex', booktabs = TRUE)
```

@@

\pagebreak
<!--- # References {#references .unnumbered} --->
# References
